---
title: "Chapter 3"
author: "Tuomas Mäkelä"
date: "16 11 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
# Doing logistic regression analysis

## 1. The data.

Just for fun and games we will print the summary of the data we are about to use.  The summary is somewhat difficult to read, so feel free to skip to the actual work.

Looking at the data:
```{r}
Data<-read.csv("DATA/Alcohol.csv")
summary(Data)
colnames(Data)
```
The data consists of 36 variables and 382 observations. The first column X is only the identification number for the data. The data is gathered from two schools with a questionnaire to students and  school reports and wrangled so that only students who studied both mathematics and portuguese are included.

We are interested in the link between alcohol consumption  (specifically binary response, high or low) and some other interesting variables. Specifically, we will look at the variables "absences", "health", "sex" and "G3", which means the mean final grade of math and portuguese of the student in question.

The basic assumptions are clear: high alcohol consumption could have an increasing effect on absences, a decreasing effect on health, sex could well have an effect on consumption (either way, but mostly women tend to drink less) and finally, the final grade should sink if one consumes a lot.

## 2. Graphical and numerical analysis
```{r}
library(dplyr)
library(ggplot2)

Data<-read.csv("DATA/Alcohol.csv")
g1 <- ggplot(data = Data, aes(x = alc_use, fill=sex))

# define the plot as a bar plot and draw it
g1 + geom_bar()


# initialize a plot of 'high_use'
g2 <- ggplot(data = Data, aes(x=high_use, fill=sex))

# draw a bar plot of high_use by sex
g2+geom_bar()

```
  
Looking at the distribution of the sex variable, we see what was assumed: women consume slightly less. Now let's draw a more detailed plot of what's going on in the variables.
  
```{r}
library(GGally)
library(ggplot2)
Data1<-select(Data, one_of(c("sex", "G3", "health", "absences", "alc_use")))
p <- ggpairs(Data1, mapping = aes(col=sex, alpha=0.3), lower = list(combo = wrap("facethist", bins = 20)))
p
```
 
 From the graphs we see that the assumptions made previously seem somewhat correct; the correlation between alc_use and G3 is negative, especially in men, while the correlation between absences and alc_use is positive. Health does not seem to have a significant relation to alc_use, probably because the effects of alcohol use do not onset until later. Interestingly, all the distributions (on the diagonal) of the variables seem to be similar fo the sexes, except for alcohol consumption, which is significantly more skewed to the left for women. This leads to the question of validity: do the students answer differently based on sex?
 
## 3. The logit model

We will now fit a logistic model to the data.
```{r}
Data<-read.csv("DATA/Alcohol.csv")
library(dplyr)

model <- glm(high_use ~ G3 + absences + health + sex, data = Data, family = "binomial")

# compute odds ratios (OR)
OR <- coef(model) %>% exp

# compute confidence intervals (CI)
CI<-exp(confint(model))

# print out the odds ratios with their confidence intervals
cbind(OR, CI)

```
Here we see the summary of the fitted logistic regression model. The odds ratio for G3 is 0.9 with the confidence interval not overlapping 1 with 95% confidence. This means that odds for high G3 values mean lower odds for alcohol consumption. For absences, the odds ratio is a little over 1, wiht the confidence interval again not containing 1. This means that statistically the higher the absences, the higher the consumption of alcoho. For health we see what previously was seen; the confidence interval contains 1, which means that statistically we cannot be sure if health has any effect, and last but _not least_, the odds ratio for males compared to odds ratio of high alcohol consumption seems to be the highest in this model. This does not mean that sex is the best bet for understanding high alcohol consumption. It only states that the probability of being male seems to give some information on alcohol consumption.

Further analysis is needed if one wants to find out the best fit model for this data.

## 4. The goodness of the model

Looking at our model we decide that since health was not seemingly statistically significant, we leave it out of the model and test the predictive power of the model

\begin{equation*}
log(\frac{p}{1-p})=\alpha +\beta_1 x_{G3} +\beta_2 x_{Abs} +\beta_3 x_{Sex} + \epsilon
\end{equation*}

Where the coefficients correspond to the normal odds ratio and the odds ratios of G3, Abs and Sex, respectively.

```{r}
Data<-read.csv("DATA/Alcohol.csv")
library(dplyr)
library(GGally)
library(ggplot2)

M<- glm(high_use ~ G3 + absences + sex, data = Data, family = "binomial")

# predict() the probability of high_use
probabilities <- predict(M, type = "response")

# add the predicted probabilities to 'alc'
Data <- mutate(Data, probability = probabilities)

# use the probabilities to make a prediction of high_use
Data <- mutate(Data, prediction = probability>0.5)

# see the last ten original classes, predicted probabilities, and class predictions
select(Data, failures, absences, sex, high_use, probability, prediction) %>% tail(10)

# tabulate the target variable versus the predictions
table(high_use = Data$high_use, prediction = Data$prediction)

g <- ggplot(Data, aes(x = high_use, y = probability, col=prediction))

# define the geom as points and draw the plot
g + geom_point()

# tabulate the target variable versus the predictions
table(high_use = Data$high_use, prediction = Data$prediction)

```

