# Regression and model validation {#anchor}

The second week of the course has taken me through some data wrangling and later a little of analysis, with emphasis on graphical analysis.  

## Data {#Data}

```{r}
Learning<-read.csv("LRN14.csv")
dim(Learning)
str(Learning)
max(Learning$Points)
min(Learning$Points)
```
Here we see some aspecsts of the data we start to work on. The data is gathered in 2014 from an introductory course on social statistics. The data consists of 7 variables and 166 observations, with the first variable being only the subject id. The variables are named so that one knows what one is dealing with for the most part, but for the variables Attitude, Deep, stra, surf and points a little clarification is necessary.

Attitude, Deep, stra and surf are all combined from several questionnaire variables with ranges 1-5, scaled so that the range of the sum is also 1-5 (by taking the mean). Points is the score of the final exam, ranging from  7 to 33, with maximum points 
Attitude is quite clear, Deep means deep, stra and surf the deep, strategid and surface type of approac to learning in the course. 

## Graphical analysis {#Ganal}
```{r echo=F}
library(GGally)
library(ggplot2)

# create a more advanced plot matrix with ggpairs()
p <- ggpairs(Learning, mapping = aes(col=gender, alpha=0.3), lower = list(combo = wrap("facethist", bins = 20)))
p

```

Here we see a detailed graphical representation of the data. Some interesting facts pop up, namely most of the subjects are female. The ranges and the distributions of the variables are evident, and the most interesting fact about them is that the distributions of the attitude and surf variables differ between males and females. Some of the correlations between the variables seem possibly significant, but one cannot really say based on this analysis. 

## Statistical analysis {#Sanal}

For a more detailed analysis, we fir a linear regression model onto the data. In the model we assume that the relationship of explanatory variables (here Attitude, stra and surf) explain the variable of interest (Points) in a linear manner. 

```{r}
library(ggplot2)
library(GGally)

# create a regression model with multiple explanatory variables
my_model2 <- lm(Points ~ Attitude + stra + surf, data = Learning)

# print out a summary of the model
summary(my_model2)
```
The statistical test (looking at the T-values and p-values) asks if the null hypothesis that the linear coefficient corresponding to the particular explanatory variable is zero. This would mean that the variable has little or no effect on the variation of the target variable, or Points in this case.

As seen from the summary, the coefficients related to stra and surf are not statistically significant and the variables should therefore be excluded from the model.

Redoing the model, this time with only the one (significant) explanatory variable:

```{r}
library(ggplot2)
library(GGally)

# create a regression model with one explanatory variable
my_model2 <- lm(Points ~ Attitude, data = Learning)

# print out a summary of the model
summary(my_model2)
```
From the renewed model we see that the relevant estimate of the coefficient is approximately 3,5. This means that Attitude toward the course explains the course grade (specifically, the total points) so that an increase of one in Attitude will lead to an increase of 3,5 in Points. The intercept parameter is the value of expected points in the exam with an attitude of expected attitude (generally the mean). 

The multiple R squared of the model is in this case simply the square of the regression between Attitude and Points. Since it is around 0,2, approximately 20% of the variation of points can be explained by the variation of Attitude.

## Model validation {#ModVal}
Now we look at our model in light of some statistical graphical representations.

```{r echo=FALSE}
library(ggplot2)
library(GGally)
my_model2 <- lm(Points ~ Attitude, data = Learning)

# draw diagnostic plots using the plot() function. Choose the plots 1, 2 and 5


plot(my_model2, which=c(1, 2, 5))
```

The assumptions of the model are that the error terms are approximately normally distributed with mean 0 and identical variation, uncorrelated and independent of the variable of interest. Specifically, the size of the error should not depend on the value of the explanatory or interesting variables.

From these pictures it seems that the model assumptions are approximately correct, with the small exception of small and large values of Points corresponding to some larger deviation from the estimated mean. Also, a couple of observations seem to have somewhat large leverages, but overall the assumptions of the model seem to hold quite well. Questionable is the ends of the spectrum, and additional analysis is needed.
